RDF Validation Requirements - Evaluation and Logical Underpinning

### 2015-01-09 paper/ESWC 2015 Tutorial/ESWC2015Tutorial.tex
typos

\begin{abstract}
The DDI-RDF Discovery Vocabulary (Disco) is an RDF Schema vocabulary that supports the discovery of microdata sets and related metadata using RDF technologies in the Web of Linked Data. Disco can be used to discover datasets by searching for specific questions, topics, and geographical coverage. Disco is intended to provide means to describe microdata by essential metadata for the discovery purpose. 
The DDI (Data Documentation Initiative) is a structured metadata standard related to the observation and measurement of human activity.

@@

This tutorial addresses Linked Data developers who want to publish Linked Data sets based on complex data like person-level microdata in combination with essential informationlike provenance, data aggregates and data catalogs.

@@

The DDI-RDF Discovery Vocabulary (Disco)\footnote{current specification is available at \url{https://github.com/linked-statistics/disco-spec}}  is an RDF Schema vocabulary that supports the discovery of microdata sets and related metadata using RDF technologies in the Web of Linked Data. Disco can be used to discover datasets by searching for specific questions, topics, and geographical coverage. Disco is intended to provide means to describe microdata by essential metadata for the discovery purpose.

@@

For a publication of DDI metadata as Linked Data, widely accepted and adopted RDF vocabularies (e.g. RDF Data Cube, DCAT, and PHDD) are reused to a large extend. It is shown how Disco is interwoven with these vocabularies. The Data Cube vocabulary is a W3C standard for representing data cubes representing multidimensional aggregate data derived from microdata which is represented by Disco. DCAT is a W3C standard for describing catalogs of datasets. Physical data description (PHDD) represents data (tables) in a rectangular format. The data could be either represented in records with character-separated values (CSV) or in records with fixed length. The combined usage of PHDD, Disco, and DCAT enables the creation of data repositories which provide metadata for the description of collections, for data discovery, and for processing of the data.

@@

\subsection{Objectives of the Tutorial} 
The domain-specific origin of Disco increases the difficulties for adopting it by developers who are not familiar with the domain. However, person-level metadata is not restricted to a particular domain and may be an interesting data source for Linked Data developers. Hence, as our main objective, we want to provide assistence for creating such complex Linked Data sets using Disco. Because of the complexity, we especially want to demonstrate how Disco data sets can be connected with information represented using other existing vocabularies that are relevant in that context. Vice verse, we show how these other vocabualries and data represented with such can benefit from the interplay with Disco.

@@

During the tutorial, we built one real world complex data set which serves as ongoing running example. 
The whole data set can be exported from the Microdata Information System\footnote{\url{http://www.gesis.org/missy/editor/}}. 
In all sessions, there will be practical hands-on exercises where the participants are asked to model and query particular parts of this data set. 
The infrastructure for these exercises, i.e. a triple store, will be provided for all participants.
We will extend slides of a previous tutorial\footnote{tutorial slides avaliable at: \url{http://de.slideshare.net/boschthomas/201412-lets-disco-eddi-2014} and \url{http://de.slideshare.net/boschthomas/201412-lets-disco-2-eddi-2014}} 
and where appropriate refer to the current Disco specification\footnote{current specification is available at \url{https://github.com/linked-statistics/disco-spec}}.

\subsection{Intended Audiences}
The intended audience of this tutorial are Linked Data developers - beginners as well as experts - who are interested in modelling and publishing complex and highly inter-connected data sets as Linked Data. Since the connection with other vocabulaires like QB, DCAT and PROV-O plays a major role in this context, the tutorial is not restricted to participants who are interested in person-level micro data.

\subsection{Learning Objectives}
The learning objectives of this tutorial are the following:
\begin{itemize} 
\item Introduction of Disco, PHDD, and XKOS for modelling complex person-level micro data
\item Representing the interplay of Disco data with existing vocabularies
\item Validation of Disco
\end{itemize}

@@

Benjamin Zapilko holds a degree in Computer Science and has recently finished his PhD thesis about methods for matching social science relevant Linked Data. The defense talk will be at the end of January. Beside his PhD research, he works on modelling and publishing social science relevant data as Linked Data with the focus on data modelling and integration of heterogeneous data sources and information types. He presented his work at several international conferences.

>>>>

\begin{abstract}
The DDI-RDF Discovery Vocabulary (Disco) is an RDF Schema vocabulary that supports the discovery of microdata sets and related metadata using RDF technologies in the Web of Linked Data. Disco can be used to discover data sets by searching for specific questions, topics, and geographical coverage. Disco is intended to provide means to describe microdata by essential metadata for the discovery purpose. 
The DDI (Data Documentation Initiative) is a structured metadata standard related to the observation and measurement of human activity.

@@

This tutorial addresses Linked Data developers who want to publish Linked Data sets based on complex data like person-level microdata in combination with essential information like provenance, data aggregates and data catalogs.

@@

The DDI-RDF Discovery Vocabulary (Disco)\footnote{current specification is available at \url{https://github.com/linked-statistics/disco-spec}}  is an RDF Schema vocabulary that supports the discovery of microdata sets and related metadata using RDF technologies in the Web of Linked Data. Disco can be used to discover data sets by searching for specific questions, topics, and geographical coverage. Disco is intended to provide means to describe microdata by essential metadata for the discovery purpose.

@@

For a publication of DDI metadata as Linked Data, widely accepted and adopted RDF vocabularies (e.g. RDF Data Cube, DCAT, and PHDD) are reused to a large extend. It is shown how Disco is interwoven with these vocabularies. The Data Cube vocabulary is a W3C standard for representing data cubes representing multidimensional aggregate data derived from microdata which is represented by Disco. DCAT is a W3C standard for describing catalogs of data sets. Physical data description (PHDD) represents data (tables) in a rectangular format. The data could be either represented in records with character-separated values (CSV) or in records with fixed length. The combined usage of PHDD, Disco, and DCAT enables the creation of data repositories which provide metadata for the description of collections, for data discovery, and for processing of the data.

@@

\subsection{Objectives of the Tutorial} 
The domain-specific origin of Disco increases the difficulties for adopting it by developers who are not familiar with the domain. However, person-level metadata is not restricted to a particular domain and may be an interesting data source for Linked Data developers. Hence, as our main objective, we want to provide assistance for creating such complex Linked Data sets using Disco. Because of the complexity, we especially want to demonstrate how Disco data sets can be connected with information represented using other existing vocabularies that are relevant in that context. Vice versa, we show how these other vocabularies and data represented with such can benefit from the interplay with Disco.

@@

During the tutorial, we built one real world complex data set which serves as ongoing running example. 
The whole data set can be exported from the Microdata Information System\footnote{\url{http://www.gesis.org/missy/editor/}}. 
In all sessions, there will be practical hands-on exercises where the participants are asked to model and query particular parts of this data set. 
The infrastructure for these exercises, i.e. a triple store, will be provided for all participants.
We will extend slides of a previous tutorial\footnote{tutorial slides available at: \url{http://de.slideshare.net/boschthomas/201412-lets-disco-eddi-2014} and \url{http://de.slideshare.net/boschthomas/201412-lets-disco-2-eddi-2014}} 
and where appropriate refer to the current Disco specification\footnote{current specification is available at \url{https://github.com/linked-statistics/disco-spec}}.

\subsection{Intended Audiences}
The intended audience of this tutorial are Linked Data developers - beginners as well as experts - who are interested in modeling and publishing complex and highly inter-connected data sets as Linked Data. Since the connection with other vocabularies like QB, DCAT and PROV-O plays a major role in this context, the tutorial is not restricted to participants who are interested in person-level micro data.

\subsection{Learning Objectives}
The learning objectives of this tutorial are the following:
\begin{itemize} 
\item Introduction of Disco, PHDD, and XKOS for modeling complex person-level micro data
\item Representing the interplay of Disco data with existing vocabularies
\item Validation of Disco
\end{itemize}

@@

Benjamin Zapilko holds a degree in Computer Science and has recently finished his PhD thesis about methods for matching social science relevant Linked Data. The defense talk will be at the end of January. Beside his PhD research, he works on modeling and publishing social science relevant data as Linked Data with the focus on data modeling and integration of heterogeneous data sources and information types. He presented his work at several international conferences.

### 2015-01-09 paper/RDF validation requirements/RDF validation requirements.tex
typos

In this paper, we describe each requirement within the RDF validation requirements database in detail (sections 3-76).
Each requirement corresponds to an RDF constraint which may be expressible by multiple constraint languages.
For each requirement, we represent some examples in different constraint languages.

@@

\begin{itemize}
	\item R-194-PROVIDE-STRING-FUNCTIONS-FOR-RDF-LITERALS
\end{itemize}

\begin{itemize}
	\item Some constraints require building new strings out of other strings
\end{itemize}

\begin{itemize}
	\item string length
\end{itemize}

@@

\begin{itemize}
	\item R-209-VALID-CLASSES
\end{itemize}

\begin{itemize}
	\item What types of resources (rdf:type) are valid in this context? Context can be an input stream, a data creation function, or an API.
\end{itemize}

@@

\begin{itemize}
	\item R-210-VALID-PROPERTIES
\end{itemize}

\begin{itemize}
	\item What properties can be used within this context? Context can be an data receipt function, data creation function, API, etc.
\end{itemize}

@@

\begin{itemize}
	\item R-31-DEFAULT-VALUES-OF-RDF-OBJECTS
  \item R-38-DEFAULT-VALUES-OF-RDF-LITERALS
\end{itemize}

\begin{itemize}
	\item default values for RDF literals
	\item default values for RDF objects
	\item default values are inferred automatically
	\item It should be possible to declare the default value for a given property, e.g. so that input forms can be pre-populated and to insert a required property that is missing in a web service call. 
\end{itemize}

@@

\begin{ex}
# rule (SPIN)
# -----------
owl:Thing
    spin:rule [
        a sp:Construct ;
            sp:text """
                CONSTRUCT {            
                    ?this :laserSwordColor "blue"^^xsd:string ;
                    ?this :numberLaserSwords "1"^^xsd:nonNegativeInteger . 
                }
                WHERE {             
                    ?this a :Jedi .            
                } """ ; ] .
owl:Thing
    spin:rule [
        a sp:Construct ;
            sp:text """
                CONSTRUCT {
                    ?this :laserSwordColor "red"^^xsd:string ;
                    ?this :numberLaserSwords "2"^^xsd:nonNegativeInteger . 
                }
                WHERE {             
                    ?this a :Sith .            
                } """ ; ] .
\end{ex}

\begin{ex}
# data
# ----
:Joda a :Jedi .
:DarthSidious a :Sith .
\end{ex}

\begin{ex}
# inferred triples
# ----------------
:Joda 
    :laserSwordColor "blue"^^xsd:string ;
    :numberLaserSwords "1"^^xsd:nonNegativeInteger .
:DarthSidious 
    :laserSwordColor "red"^^xsd:string ;
    :numberLaserSwords "2"^^xsd:nonNegativeInteger .
\end{ex}

@@

\begin{itemize}
	\item R-42-MATHEMATICAL-OPERATIONS
	\item R-41-STATISTICAL-COMPUTATIONS
\end{itemize}

\begin{itemize}
	\item adding 2 dates
	\item add number of days to start date
	\item area = width * height
	\item Statistical Computations: average, mean, sum
\end{itemize}

@@

\begin{itemize}
	\item R-47-LANGUAGE-TAG-MATCHING
\end{itemize}

@@

\begin{itemize}
	\item R-49-RDF-LITERALS-HAVING-AT-MOST-ONE-LANGUAGE-TAG
	\item R-48-MISSING-LANGUAGE-TAGS
\end{itemize}

@@

check that no language is used more than once per property

\begin{ex}
# DQTP
# ----
SELECT DISTINCT ?s WHERE { ?s %%P1%% ?c
    BIND ( lang(?c) AS ?l )
    FILTER (isLiteral (?c) && lang(?c) = %%V1%%)}
GROUP BY ?s HAVING COUNT (?l) > 1
\end{ex}

\begin{itemize}
	\item ONELANGPattern \cite{Kontokostas2014} 
  \item A literal value should contain at most 1 literal for a language. 
  \item P1 is the property containing the literal and V1 is the language we want to check.
\end{itemize}

\begin{ex}
# test binding (DQTP)
---------------------
P1 => foaf:name
V1 => en
\end{ex}

\begin{itemize}
	\item a single English (“en”) foaf:name
\end{itemize}

\begin{ex}
# valid data
# ----------
:LeiaSkywalker
    foaf:name 'Leia Skywalker'@en .
\end{ex}

\begin{ex}
# invalid data
# ------------
:LeiaSkywalker
    foaf:name 'Leia Skywalker'@en ;
    foaf:name 'Leia'@en .
\end{ex}

>>>>

In this paper, we describe each requirement within the RDF validation requirements database in detail (sections 3-76).
Additional descriptions can be found directly in the database.
Each requirement corresponds to an RDF constraint which may be expressible by multiple constraint languages.
For each requirement, we represent some examples in different constraint languages.

@@

\ms{String Operations} corresponds to the requirement
\ms{R-194-PROVIDE-STRING-} \ms{FUNCTIONS-FOR-RDF-LITERALS}.
Some constraints require building new strings out of other strings.
Calculating the string length would also be another constraint of this type.

@@

\ms{Context-Specific Valid Classes} corresponds to the requirements \\
\ms{R-209-VALID-CLASSES}.
What types of resources (rdf:type) are valid in a specific context? 
Context can be an input stream, a data creation function, or an API.

@@

\ms{Context-Specific Valid Properties} corresponds to the requirement
\ms{R-210-} \ms{VALID-PROPERTIES}.
What properties can be used within this context? 
Context can be an data receipt function, data creation function, or API.

@@

\ms{Default Values} corresponds to the requirements
\ms{R-31-DEFAULT-VALUES-OF-RDF-} \ms{OBJECTS} and
\ms{R-38-DEFAULT-VALUES-OF-RDF-LITERALS}.
Default values for objects and literals are inferred automatically.
It should be possible to declare the default value for a given property, e.g. so that input forms can be pre-populated and to insert a required property that is missing in a web service call.

@@

\begin{ex}
# rule (SPIN)
# -----------
owl:Thing
    spin:rule [
        a sp:Construct ;
            sp:text """
                CONSTRUCT {            
                    ?this laserSwordColor "blue"^^xsd:string ;
                    ?this numberLaserSwords "1"^^xsd:nonNegativeInteger . 
                }
                WHERE {             
                    ?this a Jedi .            
                } """ ; ] .
owl:Thing
    spin:rule [
        a sp:Construct ;
            sp:text """
                CONSTRUCT {
                    ?this laserSwordColor "red"^^xsd:string ;
                    ?this numberLaserSwords "2"^^xsd:nonNegativeInteger . 
                }
                WHERE {             
                    ?this a Sith .            
                } """ ; ] .
\end{ex}

\begin{ex}
# data:
Joda a Jedi .
DarthSidious a Sith .
\end{ex}

\begin{ex}
# inferred triples:
Joda 
    laserSwordColor "blue"^^xsd:string ;
    numberLaserSwords "1"^^xsd:nonNegativeInteger .
DarthSidious 
    laserSwordColor "red"^^xsd:string ;
    numberLaserSwords "2"^^xsd:nonNegativeInteger .
\end{ex}

@@

\ms{Mathematical Operations} corresponds to the requirements
\begin{itemize}
	\item \ms{R-42-MATHEMATICAL-OPERATIONS} and
  \item \ms{R-41-STATISTICAL-COMPUTATIONS}.
\end{itemize}
Examples are:
\begin{itemize}
	\item adding 2 dates
	\item add number of days to start date
	\item area = width * height
	\item Statistical Computations: average, mean, sum
\end{itemize}

@@

\ms{Language Tag Matching} corresponds to the requirement \\
\ms{R-47-LANGUAGE-TAG-MATCHING}.

@@

\ms{Language Tag Cardinality} corresponds to the requirements
\begin{itemize}
	\item \ms{R-49-RDF-LITERALS-HAVING-AT-MOST-ONE-LANGUAGE-TAG} and
	\item \ms{R-48-MISSING-LANGUAGE-TAGS}.
\end{itemize}

@@

Check that no language is used more than once per property

\begin{ex}
# DQTP:
SELECT DISTINCT ?s WHERE { ?s %%P1%% ?c
    BIND ( lang(?c) AS ?l )
    FILTER (isLiteral (?c) && lang(?c) = %%V1%%)}
GROUP BY ?s HAVING COUNT (?l) > 1
\end{ex}

This corresponds to the test pattern \ms{ONELANGPattern} \cite{Kontokostas2014}.
A literal value should contain at most 1 literal for a language. 
\ms{P1} is the property containing the literal and \ms{V1} is the language we want to check.

\begin{ex}
# test binding (DQTP):
P1 => foaf:name
V1 => en
\end{ex}

A single English (“en”) \ms{foaf:name}.

\begin{ex}
# valid data:
:LeiaSkywalker
    foaf:name 'Leia Skywalker'@en .
\end{ex}

\begin{ex}
# invalid data:
:LeiaSkywalker
    foaf:name 'Leia Skywalker'@en ;
    foaf:name 'Leia'@en .
\end{ex}
